### 2018-04-11 paper/introduction.tex

% intro:
\section{Introduction}
\label{sec:intro}
% context, problem defn:
Many critical problems in program verification can be reduced to
solving systems of Constrained Horn Clauses (CHCs), a class of
logic-programming
problems~\cite{bjorner13,flanagan03,rummer13a,rummer13b}.
%
A CHC is a logical implication where the antecedent is a conjunction
of uninterpreted relational predicates with a fact in a background
theory and the consequent is an uninterpreted relational predicate. In
the context of a CHC we call the antecedent the body and the
consequent the head.  A CHC system is a set of CHCs.
%
The goal of a CHC solving problem is to find suitable interpretations
for each relational predicate such that each CHC is logically
consistent in isolation.

% introduce the problem of solving recursion-free systems:
In this work we focus on the subclass of CHC systems which are known
as \emph{recursion-free}. In a recursion-free CHC system, the
derivation of each relational predicate will contain no references to
itself. \QZ{\DAH{I still think we need a small example here}}
%
Recursion-free CHC systems are an important subclass for two reasons.
%
First, recursion-free systems can be used to model safety properties
for hierarchical programs~\cite{lal-qadeer15,lal-qadeer-lahiri12}
(programs with only bounded iteration and recursion).
%
Second and most importantly, the task of solving a general CHC
system can reduce to the task of solving a sequence of recursion-free
systems \QZ{reference}.

Many \QZ{more references? Or remove many.} modern solvers generate
a sequence such that each subsystem is a bounded unwinding of the
input system. These solvers either attempt to show that the solution to a
given subsystem generalizes to the original system or generate a
larger subsystem for the next iteration~\cite{bjorner13}.
%
Since such solvers inspect recursion-free systems in an iterative
loop, they are heavily reliant on the performance of the underlying
recursion-free solver.
%
% current general techniques for solving recursion-free systems
Typically, even recursion-free CHC systems are not solved directly.
Instead, they are reduced to a more specific subclass of
recursion-free CHC system.
%
These classes include those of
\emph{body disjoint} (or \emph{tree structure})
systems~\cite{heizmann10,bjorner13,mcmillan14,rummer13a,rummer13b} and
of \emph{linear} systems~\cite{albarghouthi12a}.
%
We will discuss these classes in \autoref{sec:overview} and
\autoref{sec:related-work}.
%
Such classes have proved practical because a solver can find
suitable definitions for relational predicates by issuing
\emph{interpolation queries}.
%
When a class of CHC systems can be solved by interpolation queries, we
say the class is \emph{directly solvable}.

% Time complexity
In general, solving a recursion-free CHC system whose facts are
restricted to the theory of linear integer arithmetic is
co-NEXPTIME-complete~\cite{rummer13b}.
%
In contrast, solving a directly solvable system with the same theories
is in co-NP~\cite{rummer13b}.
%
Therefore, the performance of a CHC solver is highly reliant on the
size of the directly solvable systems it produces.

% contribution of this paper: CDD systems:
The first contribution of this paper is the introduction of a novel
class of directly solvable systems that we refer to as
\emph{Clause-Dependence Disjoint} (CDD).
%
The class consists of all recursion-free CHC systems such that each
clause in the system has at most one dependency on the same relational
predicate.
%
The key characteristic of this class is that CDD systems are often
smaller than equivalent systems in other directly solvable classes,
and therefore solving recursion free systems by rewriting them in CDD
form is often less computationally expensive.

>>>>

% intro:
\section{Introduction}
\label{sec:intro}
% context, problem defn:
Many critical problems in program verification can be reduced to
solving systems of Constrained Horn Clauses (CHCs), a class of
logic-programming
problems~\cite{bjorner13,flanagan03,rummer13a,rummer13b}.
%
A CHC is a logical implication where the antecedent is a conjunction
of uninterpreted predicates with a fact in a background
theory and the consequent is an uninterpreted predicate. 
%
In the context of a CHC we call the antecedent the body and the
consequent the head.  A CHC system is a set of CHCs. 
%
The goal of a CHC solving problem is to find suitable interpretations
for each predicate such that each CHC is logically
consistent in isolation.\QZ{re-write this based on other papers}

% introduce the problem of solving recursion-free systems:
In this work we focus on the subclass of CHC systems which are known
as \emph{recursion-free}. In a recursion-free CHC system, the
derivation of each predicate will not contain
itself. \QZ{\DAH{I still think we need a small example here}}
%
Recursion-free CHC systems are an important subclass for two reasons.
%
First, recursion-free systems can be used to model safety properties
for hierarchical programs~\cite{lal-qadeer15,lal-qadeer-lahiri12}
(programs with only bounded iteration and recursion).

%
Second and most importantly, A well-known approach for solving a general CHC
system can reduce to solving a sequence of recursion-free
systems \QZ{reference}.
%
Many solvers attempt to synthesize the solution of the original system 
by using the solutions of constructed recruion-free system~\cite{bjorner13}.
%
Because of this, the performance of solving general CHC system is heavily reliant on the 
performance of the solving recursion-free CHC system.
%

% current general techniques for solving recursion-free systems
Typically, recursion-free CHC systems are not solved directly.
%
Instead, they are reduced to a more specific subclass of
recursion-free CHC system.
%
These classes include those of
\emph{body disjoint} (or \emph{derivation tree})
systems~\cite{heizmann10,bjorner13,mcmillan14,rummer13a,rummer13b} and
of \emph{linear} systems~\cite{albarghouthi12a}.
%
We will discuss these classes in \autoref{sec:overview} and
\autoref{sec:related-work}.
%
Such classes can be solved by issuing
\emph{interpolation queries} for uninterpreted predicate and 
each constrains shows only one time in the interpolation query.
%
We
refer such class as \emph{directly solvable}.

% Time complexity
In general, solving a recursion-free CHC system for 
propositional logic and theory of linear integer arithmetic is
co-NEXPTIME-complete~\cite{rummer13b}.
%
In contrast, solving a directly solvable system with the same logic and theories
is in co-NP~\cite{rummer13b}.
%
Therefore, the performance of solving a recurion-free system is highly reliant on the
size of the directly solvable systems it is reduced to.

% contribution of this paper: CDD systems:
The first contribution of this paper is the introduction of a novel
class of directly solvable systems that we refer to as
\emph{Clause-Dependence Disjoint} (CDD).
%
The class consists of all recursion-free CHC systems such that each
clause in the system has at most one dependency on the same relational
predicate.\QZ{repharse this}
%
The key characteristic of this class is that CDD systems are often
smaller than equivalent systems in other directly solvable classes,
and therefore solving recursion free systems by rewriting them in CDD
form is often less computationally expensive.

### 2018-04-11 paper/related-work.tex

% interprocedural verification:
Previous work also proposed a verifier for recursive
programs~\cite{heizmann10}.
%
The proposed approach selects interprocedural paths of a program and
synthesizes invariants for each as nested interpolants.
%
Such an approach corresponds to attempting to solve a recursive CHC
system $\mathcal{S}$ by selecting derivation trees of $\mathcal{S}$
and solving each tree.

% solving recursive systems:
Previous work has proposed solvers for recursive systems that, given a
system $\mathcal{S}$, attempt to solve $\mathcal{S}$ by generating and
solving a series of recursion-free unwindings of $\mathcal{S}$.
%
In particular, \duality attempts to solve each unwinding
$\mathcal{S}'$ by selecting and solving derivation-trees of
$\mathcal{S}'$~\cite{bjorner13}.
%
Other optimizations selects derivation trees to solve using symbolic
symbolic analogs of Prolog evaluation with
tabling~\cite{jaffar09,mcmillan14}.
%
\eldarica attempts to solve each recursion-free CHC system by lazily
copying its subsystems to form a body-disjoint
over-approximation~\cite{rummer13a,rummer13b}.

% DAG interpolation:
\whale attempts to verify a sequential recursive program by generating
and solving hierarchical programs (i.e., programs that may contain
conditional branches and procedure calls, but do not contain loops or
recursion), which correspond to recursion-free CHC
systems~\cite{albarghouthi12b}.
%
To solve a particular recursion-free system $\mathcal{S}$, \whale
generates a linear inlining $\mathcal{S}'$ of $\mathcal{S}$ and solves
it using a procedure \vinta~\cite{albarghouthi12a}.
%
In general, $\mathcal{S}'$ may have size exponential in the size of
$\mathcal{S}$.

% compare sys to all of the previous work:
\sys is similar to all of the approaches given above for solving
recursion-free CHC systems in that it reduces the problem of solving a
given recursion-free CHC system $\mathcal{S}$ to solving a CHC system
in a directly-solvable class.
%
\sys is distinct from all of the approaches given above in that it
reduces solving a recursion-free CHC system to solving a
Clause-Dependent Disjoint (CDD) system.
%
CDD systems strictly contain the union of all classes of
directly-solvable CHC systems used by the above approaches, and can
themselves be solved directly.

@@

% DAG inlining:
DAG inlining, given a hierarchical program $P$, attempts to generate a
compact verification condition for $P$~\cite{lal-qadeer15}.
%
\sys contains a procedure that, given a recursion-free Horn Clause
system $\mathcal{S}$, attempts to construct a compact CDD system
$\mathcal{S}'$ such that each solution of $\mathcal{S}'$ defines a
solution of $\mathcal{S}$.
%
Because hierarchical programs and recursion-free Horn Clauses
correspond closely to each other, algorithms that operate on
hierarchical programs directly correspond to algorithms that operate
on recursion-free Horn Clauses.
%
However, an algorithm for constructing a verification condition of
hierarchical programs cannot apparently be directly used to synthesize
a solution of a recursion-free CHC system.

>>>>

% interprocedural verification:
Previous work also proposed a verifier for recursive
programs~\cite{heizmann10}.
%
The proposed approach selects interprocedural paths of a program and
synthesizes invariants for each as nested interpolants.
%
Such an approach corresponds to attempting to solve a recursive CHC
system $S$ by selecting derivation trees of $S$
and solving each tree.

% solving recursive systems:
Previous work has proposed solvers for recursive systems that, given a
system $S$, attempt to solve $S$ by generating and
solving a series of recursion-free unwindings of $S$.

%
In particular, \eldarica attempts to solve each unwinding
$S'$ by reducing to and solving body-disjoint systems~\cite{rummer13a,rummer13b}.
%

%
\duality attempts to avoid solving all 
derivation-trees (i.e body-disjoint system) by using lazy annotation~\cite{bjorner13}.
%
Other optimizations selects derivation trees to solve using symbolic
symbolic analogs of Prolog evaluation with
tabling~\cite{jaffar09,mcmillan14}.
%

\whale attempts to verify a sequential recursive program by generating
and solving hierarchical programs (i.e., programs that may contain
conditional branches and procedure calls, but do not contain loops or
recursion), which correspond to recursion-free CHC
systems~\cite{albarghouthi12b}.
%
To solve a particular recursion-free system $S$, \whale
generates a linear inlining $S'$ of $S$ and solves
it using a procedure \vinta~\cite{albarghouthi12a}.
%
In general, $S'$ may have size exponential in the size of
$S$.

% compare sys to all of the previous work:
\sys is similar to all of the approaches given above for solving 
recursion-free CHC systems in that it reduces the problem of solving a
given recursion-free CHC system $S$ to solving a CHC system
in a directly-solvable class.
%

\sys is distinct from all of the approaches given above in that it
reduces solving a recursion-free CHC system to solving a
Clause-Dependent Disjoint (CDD) system.
%
CDD systems strictly contain the union of all classes of
directly-solvable CHC systems used by the above approaches, and can
themselves be solved directly.

%
For solving genreal CHC system, \sys uses the same strategy that proposed
by above approaches, that solving a series recursion-free CHC system that unwinding 
from the original CHC system, and try to synthesize the solution for the original 
CHC system by using solutions of recursion-free CHC systems.

@@

% DAG inlining:
DAG inlining, given a hierarchical program $P$, attempts to generate a
compact verification condition for $P$~\cite{lal-qadeer15}.
%
\sys contains a procedure that, given a recursion-free Horn Clause
system $S$, attempts to construct a compact CDD system
$S'$ such that each solution of $S'$ defines a
solution of $S$.
%
Because hierarchical programs and recursion-free Horn Clauses
correspond closely to each other, algorithms that operate on
hierarchical programs directly correspond to algorithms that operate
on recursion-free Horn Clauses.
%
However, an algorithm for constructing a verification condition of
hierarchical programs cannot apparently be directly used to synthesize
a solution of a recursion-free CHC system.
